---
title: "Revised SDF Results"
author: "Ethan Roy"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sjPlot)
library(cocor)
library(sjmisc)
library(lme4)
library(MuMIn)
library(lmerTest)
require(data.table)
require(jtools)
library(nonpar)
library(lavaan)
library(psych)
library(emmeans)
library(lsr)
library(effectsize)
library(corrplot)
library(ggpubr)
library(boot)
library(patchwork)
library(mgcv)
library(purrr)
library(lmtest)
library(numbers)

select<-dplyr::select
library(r2glmm)

# Helper function for pulling out individual effect sizes
effect_size_calc = function(mod){
  
  anova_df = data.frame(anova(mod))
  effect_df = r2beta(mod, partial = TRUE)
  effect_df = data.frame(effect_df) 
  
  return(effect_df)
  
}

# Helper Function for Generating Split-Half Reliability
gen_cor = function(df, cutoff) {
  
  # Original correlation without bootstrapping
  original_corr = df %>% 
    filter(trial_onset/1000 < cutoff) %>% 
    select(pid,timepoint,firstButtonRT, answered_correctly) %>% 
    group_by(pid) %>% 
    mutate(set = mod(row_number(),2))%>% 
    group_by(pid,set) %>% 
    dplyr::summarise(count = n(), 
              sum_corr = sum(answered_correctly, na.rm = T),
              mean_rt = mean(firstButtonRT, na.rm = T), 
              rcs = (1000*sum(answered_correctly, na.rm = T))/sum(firstButtonRT, na.rm = T)) %>% 
    ungroup() %>% 
    select(pid, sum_corr,set) %>% 
    pivot_wider(names_from = 'set',values_from = 'sum_corr',names_prefix = 'set_')%>% 
    select(set_0,set_1)%>% 
    cor(use='pairwise.complete.obs')
  
  # perform spearman-brown adjustment on correlation
  spearman_brown_corr = (2*original_corr[1, 2])/(1+original_corr[1, 2])
    
  result = list(
    original_corr = spearman_brown_corr
  )
  
  return(result)
}

# Helper Function for Generating Criterion Validity (Prediciting SBAC)
gen_val = function(df, cutoff) {
  
  # Original ICC without bootstrapping
  mod_data = df %>% 
      filter(trial_onset/1000<cutoff) %>%
    group_by(pid,timepoint,sbac) %>%
    dplyr::summarise(count = n(),
              sum_corr = sum(answered_correctly, na.rm = T),
              mean_rt = mean(rt, na.rm = T),
              rcs = (60*sum(answered_correctly, na.rm = T))/sum(firstButtonRT, na.rm = T)) %>%
    ungroup() %>% 
    mutate(timepoint = timepoint)
  
    # Baseline model
    m1 <- lmer(sbac ~ timepoint + (1 | pid) ,
               data = mod_data, REML = F)
    summary_m1 <- summary(m1)
    varcomp_m1 <- VarCorr(m1)
    varcomp_m1 <- as.data.frame(varcomp_m1, comp = "Variance")[1,4]

    # Model 2: Add math fluency rcs
    m2a <- lmer(sbac ~ timepoint + rcs + (1 | pid),
               data = mod_data, REML = F)
    
    # Calculate Change Variance explained by rcs
    summary_m2a <- summary(m2a)
    varcomp_m2a <- VarCorr(m2a)
    varcomp_m2a <- as.data.frame(varcomp_m2a, comp = "Variance")[1,4]

    r2_m2a <- (varcomp_m1 - varcomp_m2a) / varcomp_m1
  
  result = list(
    original_val = r2_m2a
  )
  
  return(result)
}

```


# Load and Merge Data

```{r, include=F}

## if this isn't present, run math_fluency_cleaning.R
full_df = read_csv('/Users/Ethan/Documents/Stanford/EdNeuro/UCSF/most updated data/sea_math_fluency_cleaned.csv')

# format timepoint and make sure we only have single digit responses
full_df = full_df %>% mutate(firstButtonRT=firstButtonRT/1000,
                             timepoint = case_when(timepoint == "t0_Fall2016" ~ 0,
                               timepoint == "t1_Spring2017" ~ 1,
                               timepoint == "t2_Fall2017" ~ 2,
                               timepoint == "t3_Spring2018" ~ 3)) %>% 
  filter(correct_response<10 & correct_response>=0) 


# Load item level information
#  A tibble: 105 × 4
#    item  fluency_type class    operation  
#    <chr> <chr>        <chr>    <chr>      
#  1 3+3=  Rule         Tie      Addition   
#  2 0+1=  Rule         One_Op   Addition   
#  3 0+4=  Rule         Identity Addition  
item_info_df = read_csv('/Users/Ethan/Documents/Stanford/EdNeuro/UCSF/Paper_1_Methods/item_info.csv',show_col_types = FALSE)

# Add item level information to full dataframe
full_df = full_df %>% 
  mutate(item = paste0(gsub(' ', '', question_text),'=')) %>% 
  full_join(item_info_df %>% 
              select(item,class),
            by='item')  %>%
  mutate(class = case_when(class=='One_Op'~'OneOp', T~class)) 

# Only select the columns we want
full_df = full_df %>%
  select(pid,timepoint,question_text,firstButtonRT,answered_correctly,
         module,class,operation_type,operand_distance,set_size,cohort,
         correct_response,grade,age_years,trial_onset,rt) %>% 
  mutate(op_dist_bin = case_when(operand_distance >=5~'Large',
                                 operand_distance < 5~'Small')) 

# Load demographic and EF information
low_income_df = read.csv('low_income_df.csv')

# Load information about student school/classroom membership
class_info_df = read.csv('/Users/Ethan/Documents/Stanford/EdNeuro/UCSF/most updated data/student_class_info.csv')

## Demographic/Academic Data
# Load EF and SBAC data
abcd_or <- read_csv("/Users/Ethan/Documents/Stanford/EdNeuro/UCSF/data-selected/criticalValuesALL_averaged_data_consentcheck_2019-02-01 (1) (1).csv",
                 n_max = 4247)

colnames(abcd_or) <- gsub(".", "_", colnames(abcd_or), fixed = TRUE)
colnames(abcd_or) <- tolower(colnames(abcd_or))

abcd <- abcd_or %>%
  dplyr::rename(math_q2_2017 = x2017_math_q2_lettergrade,
         math_q4_2017 = x2017_math_q4_lettergrade,
         math_q2_2018 = x2018_math_q2_lettergrade,
         math_q4_2018 = x20_18_math_q4_lettergrade) %>%
  dplyr::mutate(time = case_when(time_point == "T1_Fall2016" ~ 0,
                          time_point == "T2_Spring2017" ~ 1,
                          time_point == "T3_Fall2017" ~ 2,
                          time_point == "T4_Spring2018" ~ 3),
         year = case_when(time == "0" | time == "1" ~ 1,
                          time == "2" | time == "3" ~ 2),
         math_fluency_acc_rt = (math_fluency_rt_count_correct / 3),
         grade_x = ifelse(is.na(grade_x), grade_y, grade_x),
         grade_x = ifelse(is.na(grade_x), grade, grade_x),
         sbacDiff = x2018_math_sbac_ss- x2017_math_sbac_ss) %>%
         select(pid, sex,grade_x,year,time_point,
         x2017_math_sbac_ss,
         x2018_math_sbac_ss,
         sbacDiff,
         sex,
         ethnicity,
         age_x,
         math_q2_2017,
         math_q4_2017,
         math_q2_2018,
         math_q4_2018,
         language_fluency,
         flanker_rcs_overall,
         low_income,
         parent_ed_lvl,time,
         x2017_math_sbac_pl,
         x2018_math_sbac_pl,
         backwardsspatialspan_object_count_span_overall,
         stroop_rcs_congruent,
         stroop_rcs_incongruent,
         taskswitch_rcs_stay,
         taskswitch_rcs_switch,
         brt_rcs_overall,
         math_fluency_acc_rt,
         spatialspan_object_count_span_overall
         )

ef_df = abcd %>%
  mutate(timepoint = case_when(time_point == "T1_Fall2016" ~ 0,
                        time_point == "T2_Spring2017" ~ 1,
                        time_point == "T3_Fall2017" ~ 2,
                        time_point == "T4_Spring2018" ~ 3),
         stroop_overall = (stroop_rcs_congruent+stroop_rcs_incongruent)/2) %>% 
  select(pid,timepoint,brt_rcs_overall,flanker_rcs_overall,
         spatialspan_object_count_span_overall,stroop_overall)

demo_df = abcd %>% 
  select(pid,low_income,sex,parent_ed_lvl)%>%
  unique() %>% 
  drop_na() %>%
  mutate(low_income = ifelse(low_income=='Yes',1,0))

# Generate SBAC data frame w/ quartiles
sbac_df = abcd %>%
  mutate(timepoint = case_when(time_point == "T1_Fall2016" ~ 0,
                        time_point == "T2_Spring2017" ~ 1,
                        time_point == "T3_Fall2017" ~ 2,
                        time_point == "T4_Spring2018" ~ 3)) %>% 
  select(pid,timepoint,contains('math_sbac')) %>%
  unique() %>% 
  drop_na() %>% 
  inner_join(full_df %>% 
               select(pid,grade,timepoint) %>%
               unique(), 
             by=c('pid','timepoint')) %>% 
  group_by(grade,timepoint) %>% 
  mutate(sbac_quartile_year_1 = ntile(x2017_math_sbac_ss,4),
         sbac_quartile_year_2 = ntile(x2018_math_sbac_ss,4)) %>% 
  mutate(sbac_quartile = case_when(timepoint==0~sbac_quartile_year_1,
                                   timepoint==1~sbac_quartile_year_1,
                                   timepoint==2~sbac_quartile_year_2,
                                   timepoint==3~sbac_quartile_year_2),
         sbac = case_when(timepoint==0~x2017_math_sbac_ss,
                          timepoint==1~x2017_math_sbac_ss,
                          timepoint==2~x2018_math_sbac_ss,
                          timepoint==3~x2018_math_sbac_ss)) %>% 
  select(pid,grade,timepoint,sbac_quartile,sbac) %>%
  ungroup()

# dataframe for criterion validity analysis
full_df_sbac = full_df %>% 
  inner_join(sbac_df %>% 
               filter((timepoint==1)|(timepoint==3)),
             by=c('pid','timepoint','grade'))

# Create mereged dataframe for participants w SDF, Demos, and SBAC
full_df_ses = full_df %>% 
  filter(timepoint==0) %>% 
  inner_join(demo_df, by='pid') %>% 
  inner_join(sbac_df, by=c('pid','timepoint','grade'))


# set up data frame for regressions
sbac_mod_df = full_df_ses %>%
  mutate(`Problem Type` = if_else(class=='Standard','Standard','Non-Standard')) %>%
    filter(!is.na(class)) %>%
    filter(module=='MATH_FLUENCY') %>%
  filter(operand_distance!=8) %>%
  group_by(pid, timepoint,grade,cohort,`Problem Type`,sbac,sbac_quartile) %>%
  dplyr::summarise(count = n(),
            sum_corr = sum(answered_correctly, na.rm = T),
            mean_rt = mean(firstButtonRT, na.rm = T),
            rcs = (60*sum(answered_correctly, na.rm = T))/sum(firstButtonRT, na.rm = T)) %>%
  ungroup() %>%
  select(pid,timepoint,grade,cohort,`Problem Type`,sum_corr,rcs,sbac) %>%
  pivot_wider(id_cols = c(pid,timepoint,grade,cohort,sbac),
              names_from = `Problem Type`,
              values_from = c(sum_corr,rcs)) %>%
  mutate(sum_corr = `sum_corr_Non-Standard`+sum_corr_Standard) %>%
  select(pid,timepoint,grade,cohort,sum_corr,rcs_Standard,`rcs_Non-Standard`,sbac) %>%
  inner_join(ef_df, by = c('pid','timepoint'))%>%
  inner_join(demo_df, by='pid') %>%
  ungroup() %>% 
  rename(`Spatial Span` = spatialspan_object_count_span_overall,
         Flanker = flanker_rcs_overall,
         Stroop = stroop_overall
         ) %>% 
  mutate_if(is.numeric, scale)

# finalize data frames for other analyses
full_df = full_df %>% 
  dplyr::mutate(operands = lapply(str_extract_all(question_text,"\\d"),as.numeric)) %>% 
  mutate(problem_size = as.numeric(lapply(operands, max))+as.numeric(lapply(operands, min))) %>% 
  ungroup() %>% 
  mutate(second_op = sapply(operands, function(x) x[2])) %>% 
  select(-operands)%>% 
  filter(timepoint==0)

# full data including SES/SBAC info
full_df_ses = full_df_ses %>% 
 dplyr::mutate(operands = lapply(str_extract_all(question_text,"\\d"),as.numeric)) %>% 
  mutate(problem_size = as.numeric(lapply(operands, max))+as.numeric(lapply(operands, min))) %>% 
  ungroup() %>% 
  mutate(second_op = sapply(operands, function(x) x[2])) %>% 
  select(-operands) %>% 
  filter(timepoint==0)

```

## Examine number of trials completed

```{r, echo = FALSE, message=FALSE, warning = FALSE}

## Range of completed trials: 17-51

# get min number of trials
full_df %>% 
  group_by(pid) %>%
  summarise(count=n()) %>%
  select(count) %>% min()

# get max number of trials
full_df %>% 
  group_by(pid) %>%
  summarise(count=n()) %>%
  select(count) %>% max()

```


## Check out missing data by school

```{r, include=F}

# Generate data frame with just participants who provided demo information
parent_inc_df = abcd %>% 
  inner_join(class_info_df %>% filter(timepoint==0) %>% 
               select(pid, School,Teacher),
             by='pid') %>% 
  select(pid,low_income,sex,parent_ed_lvl, School,Teacher)%>%
  unique() %>% 
  mutate(missing = is.na(low_income)) %>%  # use this as a proxy -- all other demo info missing too
  select(pid, missing, School,Teacher)

# Get information about SDF completion and demo data
missing_df = full_df %>% 
  inner_join(parent_inc_df, by = 'pid') %>%
  select(pid,missing,cohort,Teacher, School) %>%
  unique() 

# n = 824 with all data
missing_df %>% 
  filter(missing==F) %>% 
  select(pid) %>% 
  unique()

# Calculate the proportion of participants at each school that are missing
# demographic information
missing_df %>% 
  group_by(School) %>% 
  mutate(count=n()) %>% 
  ungroup() %>% 
  group_by(missing, School) %>% 
  summarise(prop = n()/count) %>% 
  unique()

```
# Results

## Evaluating the Reliability and Validity of the Single-Digit Fluency Assessment

### Split-Half Reliability based on amount of data
```{r, include=F}

Seconds = seq(15, 180, by=15)
splithalf_rel = data.frame(time=double(),
                           corr=double(),
                           boot_corr=double(),
                           LCI=double(),
                           UCI=double(),
                           stringsAsFactors=T)

# Loop through temporal cutoff points and calculate split-half reliability
for (time in Seconds){
  
  split_half = gen_cor(full_df,time)
  
  res = as.data.frame(list(time=time, corr=split_half$original_corr))
  
  splithalf_rel = rbind(splithalf_rel,res)
  
}

## Find point at which reliability does not increase by more than 1%
d1 <- diff(splithalf_rel$corr)
thresh_idx = which.min(d1 > 0.01) 

thresh = splithalf_rel$time[thresh_idx]

splithalf_plot = splithalf_rel %>% 
  ggplot(aes(x=time,y=corr))+
  geom_point(color='blue')+
  geom_line(color='blue')+
  geom_vline(xintercept=thresh,color='black',linetype=2)+
  # geom_errorbar(aes(ymin=LCI, ymax=UCI),width=0.9, color='black')+
  theme_classic()+
  coord_cartesian(ylim = c(0.7, 1.00),
                  xlim=c(0,200))+
  ylab("Split-Half Reliability (Pearson's r)")+
  xlab("Seconds")

```

### Predicting SBAC based on amount of data
```{r, include=F}
## Validity -- Predicting SBAC scores -- One minute
Seconds = seq(15, 180, by=15)
r2_df = data.frame(time=double(),
                   r2=double(),
                   stringsAsFactors=T)

# Loop through temporal cutoff points and calculate change in variance for models
# predicting SBAC scores
for (time in Seconds){
  
  r2_boot_res = gen_val(full_df_sbac,
                        time)
  
  res = as.data.frame(list(time=time, r2=r2_boot_res$original_val))
  
  r2_df = rbind(r2_df,res)
  
}

## Find point at which r2 does not increase by more than 1%
d2 <- diff(r2_df$r2)
thresh_idx = which.min(d2 > 0.01) 

thresh = r2_df$time[thresh_idx]

# Generate Plot
r2_plot = r2_df %>% 
  ggplot(aes(x=time,y=r2))+
  geom_point(color='red')+
  geom_line(color='red')+
  geom_vline(xintercept=r2_df$time[8],color='black',linetype=2)+
  theme_classic()+
  labs(y = bquote(R^2), x = "Seconds")
```

## Figure 1

```{r, echo = FALSE, message=FALSE, warning = FALSE}
# Figure 1
(splithalf_plot | r2_plot)


```

# Chained-Fluency Assessment on a Tablet Replicates Known Effects from Discrete-Trials

## Set Size By Operation Effects -- Main Text
```{r setsize, echo = FALSE, message=FALSE, warning = FALSE}

# Generate set size*operation plot -- Figure 2A
setsize_plot = full_df_ses %>%
  group_by(pid, cohort, problem_size,operation_type) %>%
  dplyr::summarise(count = n(),
            sum_corr = sum(answered_correctly, na.rm = T),
            mean_rt = mean(firstButtonRT, na.rm = T),
            rcs = (60*sum(answered_correctly, na.rm = T))/sum(firstButtonRT, na.rm = T)) %>%
  ungroup() %>%
  mutate(cohort=as.factor(cohort+3+cohort)) %>% 
  ggplot(aes(x = problem_size, y = rcs,
             color=cohort,fill = cohort, group=cohort))+
  geom_smooth(method='lm')+
  ylab("Correct Responses per Minute")+
  xlab("Set Size")+
  labs(color = "Grade", fill = "Grade")+
  # ggtitle("RPM by Set Size")+
  theme_classic()+
  facet_wrap(.~operation_type, scale='free_x')+
  ylim(19,42)


# Generate dataframe for regression model
mod_data = full_df_ses %>%
  group_by(pid, cohort, problem_size,operation_type) %>%
  dplyr::summarise(count = n(),
            sum_corr = sum(answered_correctly, na.rm = T),
            mean_rt = mean(firstButtonRT, na.rm = T),
            rcs = (60*sum(answered_correctly, na.rm = T))/sum(firstButtonRT, na.rm = T)) %>%
  ungroup()

# Fit model
setsize_mod = lmer(rcs ~ problem_size*operation_type*cohort+(1|pid), data = mod_data)

# Summarize effects
summary(setsize_mod)

# Effect Sizes

effect_size_calc(setsize_mod) %>% 
  select(Effect,Rsq)

```


### Problem Type Effect -- Main Text
```{r problem_type, echo = FALSE, message=FALSE, warning = FALSE}

# Generate problem type plot -- Figure 2B
problem_type_plot_ses = full_df_ses %>%
  mutate(`Problem Type` = if_else(class=='Standard','Common','Exceptional')) %>%
  group_by(pid, grade,`Problem Type`) %>%
  dplyr::summarise(count = n(),
            sum_corr = sum(answered_correctly, na.rm = T),
            mean_rt = mean(firstButtonRT, na.rm = T),
            rcs = (60*sum(answered_correctly, na.rm = T))/sum(firstButtonRT, na.rm = T)) %>%
  ungroup() %>%
  mutate(accuracy=sum_corr/count) %>%
  group_by(grade,`Problem Type`) %>%
  summarise(count=n(),
            mean_rpm = mean(rcs, na.rm=T),
            sd_rpm = sd(rcs, na.rm=T)) %>% 
      mutate(se = sd_rpm/sqrt(count),
             lower.ci = mean_rpm - qt(1 - (0.05 / 2), count - 1) * se,
             upper.ci = mean_rpm + qt(1 - (0.05 / 2), count - 1) * se)%>%
  ggplot(aes(x = grade, y = mean_rpm))+
  geom_point(aes(color = `Problem Type`))+
  geom_line(aes(color = `Problem Type`))+
  geom_ribbon(aes(ymin=lower.ci,ymax=upper.ci,fill=`Problem Type`),alpha=0.25)+
  ylab("Correct Responses per Minute")+
  xlab("Grade")+
  labs(color = "Problem Set", fill = "Problem Set")+
  # ggtitle("RPM by Problem Type")+
  theme_classic()+
  ylim(19,42)

# Generate data frame for regression models
mod_data = full_df_ses %>%
    mutate(ProblemType = if_else(class=='Standard','Common','Exceptional')) %>%
  group_by(pid, cohort, ProblemType) %>%
  dplyr::summarise(count = n(),
            sum_corr = sum(answered_correctly, na.rm = T),
            mean_rt = mean(firstButtonRT, na.rm = T),
            rcs = (60*sum(answered_correctly, na.rm = T))/sum(firstButtonRT, na.rm = T)) %>% 
  ungroup() 

probTypeMod = lmer(rcs ~ ProblemType*cohort + (1|pid), data = mod_data)

summary(probTypeMod)


# Effect Sizes

effect_size_calc(probTypeMod) %>% 
  select(Effect,Rsq)

```

## Figure 2
```{r, echo = FALSE, message=FALSE, warning = FALSE}

# Generate Figure 2

((setsize_plot|problem_type_plot_ses)) + plot_annotation(tag_levels = 'A') & theme(legend.position = 'bottom',
                                                          legend.key.size = unit(0.6, 'cm'),
                                                          legend.text = element_text(size=13),
                                                          legend.title = element_text(size=13),
                                                          plot.tag = element_text(size = 25))

```



## Corrected p-values

```{r, echo = FALSE, message=FALSE, warning = FALSE}

## Get FDR Corrected p-values for effects

probSum = as.data.frame(summary(probTypeMod)$coefficients) %>% 
  mutate(model="Problem Type")

setSizeSum = as.data.frame(summary(setsize_mod)$coefficients) %>% 
  mutate(model="SetSize")

# Print adjusted p-values
probSum %>% 
  rbind(setSizeSum) %>% 
  mutate(p.adj_fdr = p.adjust(`Pr(>|t|)`,"fdr"),
         p.adj_bonf = p.adjust(`Pr(>|t|)`,"bonf")) %>% 
  select(model, p.adj_fdr,p.adj_bonf)


```


# RPM Fluency Measures Better Predict Math Achievement than Traditional Raw Scores

## Correlations

```{r, echo = FALSE, message=FALSE, warning = FALSE}

## Calculate and compare correlations


mod_cor = sbac_mod_df %>% select(is.numeric) %>% 
  select(c('rcs_Standard','rcs_Non-Standard','sum_corr','sbac'))%>% 
  rename( RCS_Common=rcs_Standard,
          RCS_Exceptional=`rcs_Non-Standard`,
         `Total Correct`=sum_corr)%>% 
  filter(!is.na(sbac)) 

# raw correlations
mod_cor %>% cor(use="pairwise.complete.obs")

# z-transformed correlations
corr_list = fisherz(mod_cor %>% cor(use="pairwise.complete.obs"))


# Compare two correlations based on two independet groups

print("Compare Rule/No-Rule correlations: ")
cocor.dep.groups.overlap(r.jk=cor(mod_cor$sbac,mod_cor$`RCS_Common`)[1][1], 
                         r.jh=cor(mod_cor$sbac,mod_cor$RCS_Exceptional)[1][1], 
                         r.kh=cor(mod_cor$RCS_Exceptional,mod_cor$`RCS_Common`)[1][1], 
                         n=length(unique(sbac_mod_df$pid)),
                         alternative="two.sided", alpha=0.05, conf.level=0.95, null.value=0)

print("Compare Total Correct/No-Rule correlations: ")
cocor.dep.groups.overlap(r.jk=cor(mod_cor$sbac,mod_cor$`RCS_Common`)[1][1], 
                         r.jh=cor(mod_cor$sbac,mod_cor$`Total Correct`)[1][1], 
                         r.kh=cor(mod_cor$`Total Correct`,mod_cor$`RCS_Common`)[1][1], 
                         n=length(unique(sbac_mod_df$pid)),
                         alternative="two.sided", alpha=0.05, conf.level=0.95, null.value=0)

print("Compare Rule/Total Correct correlations: ")
cocor.dep.groups.overlap(r.jk=cor(mod_cor$sbac,mod_cor$`RCS_Exceptional`)[1][1], 
                         r.jh=cor(mod_cor$sbac,mod_cor$`Total Correct`)[1][1], 
                         r.kh=cor(mod_cor$`Total Correct`,mod_cor$`RCS_Exceptional`)[1][1], 
                         n=length(unique(sbac_mod_df$pid)),
                         alternative="two.sided", alpha=0.05, conf.level=0.95, null.value=0)


```

## Regression Models -- Table 1
```{r,echo = FALSE, message=FALSE, warning = FALSE}

# convert cohort to numeric
sbac_mod_df$cohort = as.numeric(sbac_mod_df$cohort)

# Total Correct Model
mod1 = lm(sbac ~   cohort+low_income + brt_rcs_overall + Flanker + `Spatial Span`+sum_corr,
          data = sbac_mod_df)

#  Common RCS Model
mod2 = lm(sbac ~cohort+low_income + brt_rcs_overall + Flanker + `Spatial Span`+rcs_Standard, 
          data = sbac_mod_df)

#  Exceptional RCS Model
mod3 = lm(sbac ~  cohort+low_income + brt_rcs_overall + Flanker + `Spatial Span`+`rcs_Non-Standard`,
          data = sbac_mod_df)

# Common + Exceptional RCS Model 
mod4 = lm(sbac ~  cohort+low_income + brt_rcs_overall + Flanker + `Spatial Span`+`rcs_Non-Standard`+rcs_Standard 
          , data = sbac_mod_df)

anova(mod2,mod3,mod4, test = "Chisq")
# Compare non-nested models
coxtest(mod1,mod4)


# Table 1
tab_model(mod1, mod2, mod3,mod4, auto.label = T,
          p.style = "stars",
          dv.labels = c("Total Correct", "Non-Standard", "Standard", "Non-Standard+Standard"),
          show.ci = F,show.aic = T,show.loglik=T)


```


# Speed-Accuracy Tradeoffs Vary Based on Problem Type, Grade, and Math Achievement

## SAT by Problem Type Overall -- Figure 3A

```{r, include=F}

# load SAT code --- From Domingue et al (2022)
# This code is largely adapted from https://github.com/ben-domingue/rt_meta/tree/master/rtmeta_pkg/R
source("speed_accuracy_tradeoff_code.R", local = knitr::knit_global())

# Generate SAT data frame from full data and sbac dataframe
full_df_sat = full_df_ses %>% 
  # inner_join(sbac_df, by = c("pid","grade","timepoint")) %>% 
  mutate(ProblemType = if_else(class=='Standard','Common','Exceptional')) %>%
  mutate(sbac_decile = as.factor(ntile(sbac, 10)))

# Run SAT analysis on Exceptional problems
L = full_df_sat %>% 
  gen_x_sdf() %>% 
  qc() %>% 
  irt() %>% 
  interplay(nboot=100,nms=c("item","id","diff","th","pv","rt",
                         'ProblemType','cohort','sbac_quartile'),
            filter=c('ProblemType=="Exceptional"'))

# Run SAT analysis on Common problems
L2 = full_df_sat %>% 
    mutate(ProblemType = if_else(class=='Standard','Common','Exceptional')) %>%
  gen_x_sdf() %>% 
  qc() %>% 
  irt() %>% 
  interplay(nboot=100,
                   nms=c("item","id","diff","th","pv","rt",
                         'ProblemType','cohort','sbac_quartile'),
            filter=c('ProblemType=="Common"')) 

# Plot SAT curves -- Figure 3A
sat_plot = as.data.frame(L$pts) %>% 
  rename('p(corr)'=V2,
        'log(rt)'=V1) %>% 
  mutate(ProblemType='Exceptional')%>% 
  rbind(as.data.frame(L2$pts) %>% 
  rename('p(corr)'=V2,
        'log(rt)'=V1) %>% 
  mutate(ProblemType='Common')) %>% 
  mutate(rt = exp(`log(rt)`))%>% 
  ggplot(aes(x=rt,y=`p(corr)`,color=ProblemType))+
  geom_line()+
  geom_ribbon(aes(ymin=cil,ymax=cih,fill=ProblemType),
              alpha = 0.2,colour = NA)+
  theme_classic()+
  theme(legend.position = 'bottom')+
  ylab("P(Correct)")+
  xlab("Reaction Time")+
  labs(color = "Problem Set", fill = "Problem Set")


```

### Compare R2 of linear and quadratic models to assess (non-)linearity of SAT curves

```{r, include=F}

# Build dataframe from SAT results to compare fits of models predicting
# probability of a correct response from reaction time with either
# linear term or quadratic term

stats_df = as.data.frame(L$pts) %>%
  rename('p(corr)'=V2,
        'log(rt)'=V1) %>%
  mutate(ProblemType='Exceptional')%>%
  rbind(as.data.frame(L2$pts) %>%
  rename('p(corr)'=V2,
        'log(rt)'=V1) %>%
  mutate(ProblemType='Common')) %>%
  mutate(rt = exp(`log(rt)`)) %>%
  mutate(rt_2 = rt*rt)


## Compare linear vs curvilinear fits for problem types

# Linear Model for Exceptional Problems
lin_exceptional_mod = lm(`p(corr)`~rt, 
                         data = stats_df %>% filter(ProblemType=='Exceptional'))

# Quadratic Model for Exceptional Problems
quad_exceptional_mod = lm(`p(corr)`~ rt + I(rt^2), 
                          data = stats_df %>% filter(ProblemType=='Exceptional'))


# Linear Model for Common Problems
lin_common_mod = lm(`p(corr)`~rt, 
                    data = stats_df %>% filter(ProblemType=='Common'))

# Linear Model for Exceptional Problems
quad_common_mod = lm(`p(corr)`~ rt + I(rt^2),
                     data = stats_df %>% filter(ProblemType=='Common'))


# Build data frame that records the variance explained for each model, 
# as well as the change in R2 with the addition of a quadratic term
delta_r2_df = data.frame(
                  ex_lin_r2=summary(lin_exceptional_mod)$r.squared,
                  ex_quad_r2=summary(quad_exceptional_mod)$r.squared,
                  delta_ex_r2=summary(quad_exceptional_mod)$r.squared-summary(lin_exceptional_mod)$r.squared,
                  com_lin_r2=summary(lin_common_mod)$r.squared,
                  com_quad_r2=summary(quad_common_mod)$r.squared,
                  delta_com_r2=summary(quad_common_mod)$r.squared-summary(lin_common_mod)$r.squared,
                  Set = 'Full'
                  )
                
```


### SAT by Grade Cohort -- Figure 3B

```{r, include=F}

# By Cohort
cohorts = unique(full_df_ses$cohort)
## Loop through each grade cohort and calculate SAT by problem type
for(this_cohort in cohorts){
  
  print(this_cohort)
  cohort_df = full_df_sat %>% 
    filter(cohort==this_cohort)
  print(unique(cohort_df$cohort))
  ## Generate dataframes that fit with Ben's Code
  x_rule = gen_x_sdf(cohort_df)
  x_norule = gen_x_sdf(cohort_df)
  
  print(unique(x_rule$cohort))
  ## QC on data
  x_rule = qc(x_rule)
  x_norule = qc(x_norule)
  
  ## Fit IRT models
  rule_irt = irt(x_rule)
  norule_irt = irt(x_norule)
  
  filter_list_rule = c("ProblemType=='Exceptional'",paste0("cohort=='",as.character(this_cohort),"'"))
  filter_list_norule = c("ProblemType=='Common'",paste0("cohort=='",as.character(this_cohort),"'"))
  
  ## Calculate SAT model
  nms = c("item","id","diff","th","pv","rt", 'ProblemType','cohort','sbac_quartile')
  rule_inter = interplay(rule_irt,nboot=100,filter=filter_list_rule,nms=nms)
  norule_inter= interplay(norule_irt,nboot=100,filter=filter_list_norule,nms=nms)
  
  # generate plotting data frame for each cohort
  assign(paste0("plot_df_",this_cohort),
         as.data.frame(rule_inter$pts) %>% 
            rename('p(corr)'=V2,
                  'log(rt)'=V1) %>% 
            mutate(ProblemType='Exceptional') %>% 
            rbind(as.data.frame(norule_inter$pts) %>% 
            rename('p(corr)'=V2,
                  'log(rt)'=V1) %>% 
            mutate(ProblemType='Common')) %>% 
           mutate(cohort=this_cohort))
  
  # Build dataframe from SAT results to compare fits of models predicting
  # probability of a correct response from reaction time with either
  # linear term or quadratic term 
  stats_df = as.data.frame(rule_inter$pts) %>% 
  rename('p(corr)'=V2,
        'log(rt)'=V1) %>% 
  mutate(ProblemType='Exceptional')%>% 
  rbind(as.data.frame(norule_inter$pts) %>% 
  rename('p(corr)'=V2,
        'log(rt)'=V1) %>% 
  mutate(ProblemType='Common')) %>% 
  mutate(rt = exp(`log(rt)`)) %>% 
  # mutate_if(is.numeric, scale) %>%
  mutate(rt_2 = rt*rt)
  
  # Fit linear and quadratic models for common problems
  lin_common_mod = lm(`p(corr)`~rt, data = stats_df %>% filter(ProblemType=='Common'))
  quad_common_mod = lm(`p(corr)`~ rt + I(rt^2), data = stats_df %>%
                         filter(ProblemType=='Common'))
  
  # Fit linear and quadratic models for exceptional problems
  lin_exceptional_mod = lm(`p(corr)`~rt, data = stats_df %>% filter(ProblemType=='Exceptional'))
  quad_exceptional_mod = lm(`p(corr)`~ rt + I(rt^2), data = stats_df %>%
                         filter(ProblemType=='Exceptional'))
  
  # Add cohort level information about change in R2 to existing dataframe
  delta_r2_df = delta_r2_df %>% 
     rbind(data.frame(ex_lin_r2=summary(lin_exceptional_mod)$r.squared,
                         ex_quad_r2=summary(quad_exceptional_mod)$r.squared,
                         delta_ex_r2=summary(quad_exceptional_mod)$r.squared-summary(lin_exceptional_mod)$r.squared,
                         com_lin_r2=summary(lin_common_mod)$r.squared,
                         com_quad_r2=summary(quad_common_mod)$r.squared,
                         delta_com_r2=summary(quad_common_mod)$r.squared-summary(lin_common_mod)$r.squared,
                         Set = paste0('Cohort ',this_cohort)))

}

# Generate Figure 3B
cohort_sat_plot = plot_df_0 %>% 
  rbind(plot_df_1) %>% 
  rbind(plot_df_2) %>% 
  mutate(cohort = case_when(cohort==0~'Third Grade',
                            cohort==1~'Fifth Grade',
                            cohort==2~'Seventh Grade')) %>% 
  mutate(cohort = factor(cohort, levels=c('Third Grade','Fifth Grade', 'Seventh Grade'))) %>% 
  mutate(rt = exp(`log(rt)`))%>% 
  ggplot(aes(x=rt,y=`p(corr)`,color=ProblemType))+
  geom_line()+
  geom_ribbon(aes(ymin=cil,ymax=cih,fill=ProblemType),
              alpha = 0.2,colour = NA)+
  theme_classic()+
  theme(legend.position = 'bottom')+
  facet_grid(.~cohort)+
  ylab("P(Correct)")+
  xlab("Reaction Time")+
  ggtitle("Grade")+
  labs(color = "Problem Set", fill = "Problem Set")


```


### SAT by SBAC Quartile -- Figure 3C

```{r, include=F}
# By SBAC

quartiles = unique(full_df_sat$sbac_quartile)
# Loop through the SBAC Quartiles
for(this_quartile in sort(quartiles)){
  
  print(this_cohort)
  cohort_df = full_df_sat %>%
    filter(sbac_quartile==this_quartile)
  print(unique(cohort_df$sbac_quartile))
  ## Generate dataframes that fit with Ben's Code
  x_rule = gen_x_sdf(cohort_df)
  x_norule = gen_x_sdf(cohort_df)
  
  print(unique(x_rule$cohort))
  ## QC on data
  x_rule = qc(x_rule)
  x_norule = qc(x_norule)
  
  ## Fit IRT models
  rule_irt = irt(x_rule)
  norule_irt = irt(x_norule)
  
  filter_list_rule = c("ProblemType=='Exceptional'",paste0("sbac_quartile=='",as.character(this_quartile),"'"))
  
  filter_list_norule = c("ProblemType=='Common'",paste0("sbac_quartile=='",as.character(this_quartile),"'"))
  
  ## Calculate SAT model
  nms = c("item","id","diff","th","pv","rt", 'ProblemType','cohort','sbac_quartile')
  rule_inter = interplay(rule_irt,nboot=100,filter=filter_list_rule,nms=nms)
  norule_inter= interplay(norule_irt,nboot=100,filter=filter_list_norule,nms=nms)
  
  # Generate dataframe for plotting for each quartile
  assign(paste0("plot_df_",this_quartile),
         as.data.frame(rule_inter$pts) %>% 
            rename('p(corr)'=V2,
                  'log(rt)'=V1) %>% 
            mutate(ProblemType='Exceptional') %>% 
            rbind(as.data.frame(norule_inter$pts) %>% 
            rename('p(corr)'=V2,
                  'log(rt)'=V1) %>% 
            mutate(ProblemType='Common')) %>% 
           mutate(sbac_quartile=this_quartile))
  
  # Build dataframe from SAT results to compare fits of models predicting
  # probability of a correct response from reaction time with either
  # linear term or quadratic term 
  stats_df = as.data.frame(rule_inter$pts) %>% 
    rename('p(corr)'=V2,
          'log(rt)'=V1) %>% 
    mutate(ProblemType='Exceptional')%>% 
    rbind(as.data.frame(norule_inter$pts) %>% 
    rename('p(corr)'=V2,
          'log(rt)'=V1) %>% 
    mutate(ProblemType='Common')) %>% 
    mutate(rt = exp(`log(rt)`)) %>% 
    # mutate_if(is.numeric, scale) %>%
    mutate(rt_2 = rt*rt)
  
  # Fit linear and quadratic models for common problems
  lin_common_mod = lm(`p(corr)`~rt, data = stats_df %>% filter(ProblemType=='Common'))
  quad_common_mod = lm(`p(corr)`~ rt + I(rt^2), data = stats_df %>%
                         filter(ProblemType=='Common'))

  # Fit linear and quadratic models for Exceptional problems
  lin_exceptional_mod = lm(`p(corr)`~rt, data = stats_df %>% filter(ProblemType=='Exceptional'))
  quad_exceptional_mod = lm(`p(corr)`~ rt + I(rt^2), data = stats_df %>%
                         filter(ProblemType=='Exceptional'))
  
  # Add quartile level information about change in R2 to existing dataframe
  delta_r2_df = delta_r2_df %>% 
     rbind(data.frame(ex_lin_r2=summary(lin_exceptional_mod)$r.squared,
                         ex_quad_r2=summary(quad_exceptional_mod)$r.squared,
                         delta_ex_r2=summary(quad_exceptional_mod)$r.squared-summary(lin_exceptional_mod)$r.squared,
                         com_lin_r2=summary(lin_common_mod)$r.squared,
                         com_quad_r2=summary(quad_common_mod)$r.squared,
                         delta_com_r2=summary(quad_common_mod)$r.squared-summary(lin_common_mod)$r.squared,
                         Set = paste0('Quartile ', this_quartile)))
}

# Build Figure 3C
sbac_sat_plot = plot_df_1 %>% 
  rbind(plot_df_2) %>% 
  rbind(plot_df_3) %>% 
  rbind(plot_df_4) %>% 
    mutate(rt = exp(`log(rt)`),
           sbac_quartile = case_when(sbac_quartile==1~'Lowest',
                                     sbac_quartile==2~'Low',
                                     sbac_quartile==3~'High',
                                     sbac_quartile==4~'Highest'))%>% 
  mutate(sbac_quartile = factor(sbac_quartile,  
                                   levels=c("Lowest","Low","High","Highest"))) %>% 
  ggplot(aes(x=rt,y=`p(corr)`,color=ProblemType))+
  geom_line()+
  theme_classic()+
  geom_ribbon(aes(ymin=cil,ymax=cih,fill=ProblemType),
              alpha = 0.2,colour = NA)+
  facet_grid(.~sbac_quartile)+
  theme(legend.position="bottom")+
  ylab("P(Correct)")+
  xlab("Reaction Time")+
  ggtitle("SBAC Quartile")+
  labs(color = "Problem Set", fill = "Problem Set")

```


## Figure 3

```{r, echo = FALSE, message=FALSE, warning = FALSE}

## Figure 3
(sat_plot + (cohort_sat_plot/sbac_sat_plot))+
  plot_annotation(tag_levels = 'A')+
  plot_layout(guides = "collect") &
  theme(legend.position='bottom',
        plot.tag = element_text(size = 20, hjust = 0, vjust = 0))



```

### SAT analysis by SBAC decile

```{r, include=F}
delta_r2_df_dec = data.frame()
anova_df = data.frame()

deciles = unique(full_df_sat$sbac_decile)
for(this_decile in deciles){
  
  decile_df = full_df_sat%>%
    filter(sbac_decile==this_decile)

  ## Generate dataframes that fit with Ben's Code
  x_rule = gen_x_sdf(decile_df)
  x_norule = gen_x_sdf(decile_df)
  
  ## QC on data
  x_rule = qc(x_rule)
  x_norule = qc(x_norule)
  
  ## Fit IRT models
  rule_irt = irt(x_rule)
  norule_irt = irt(x_norule)
  
  filter_list_rule = c("ProblemType=='Exceptional'",paste0("sbac_decile=='",as.character(this_decile),"'"))
  filter_list_norule = c("ProblemType=='Common'",paste0("sbac_decile=='",as.character(this_decile),"'"))
  
  ## Calculate SAT model
  nms = c("item","id","diff","th","pv","rt", 'ProblemType','cohort','sbac_decile')
  rule_inter = interplay(rule_irt,nboot=100,filter=filter_list_rule,nms=nms)
  norule_inter= interplay(norule_irt,nboot=100,filter=filter_list_norule,nms=nms)
  
  assign(paste0("plot_df_",this_decile),
         as.data.frame(rule_inter$pts) %>% 
            rename('p(corr)'=V2,
                  'log(rt)'=V1) %>% 
            mutate(ProblemType='Exceptional') %>% 
            rbind(as.data.frame(norule_inter$pts) %>% 
            rename('p(corr)'=V2,
                  'log(rt)'=V1) %>% 
            mutate(ProblemType='Common')) %>% 
           mutate(sbac_decile=this_decile))
  
  stats_df = as.data.frame(rule_inter$pts) %>% 
    rename('p(corr)'=V2,
          'log(rt)'=V1) %>% 
    mutate(ProblemType='Exceptional')%>% 
    rbind(as.data.frame(norule_inter$pts) %>% 
    rename('p(corr)'=V2,
          'log(rt)'=V1) %>% 
    mutate(ProblemType='Common')) %>% 
    mutate(rt = exp(`log(rt)`)) %>% 
    # mutate_if(is.numeric, scale) %>%
    mutate(rt_2 = rt*rt)
  
  # Fit linear and quadratic models for Common problems
  lin_common_mod = lm(`p(corr)`~rt, data = stats_df %>% filter(ProblemType=='Common'))
  quad_common_mod = lm(`p(corr)`~ rt + I(rt^2), data = stats_df %>%
                         filter(ProblemType=='Common'))
  
  # Fit linear and quadratic models for Exceptional problems
  lin_exceptional_mod = lm(`p(corr)`~rt, data = stats_df %>% filter(ProblemType=='Exceptional'))
  quad_exceptional_mod = lm(`p(corr)`~ rt + I(rt^2), data = stats_df %>%
                         filter(ProblemType=='Exceptional'))
  
  
  delta_r2_df_dec = delta_r2_df_dec %>% 
     rbind(data.frame(ex_lin_r2=summary(lin_exceptional_mod)$r.squared,
                      ex_quad_r2=summary(quad_exceptional_mod)$r.squared,
                      delta_ex_r2=summary(quad_exceptional_mod)$r.squared-summary(lin_exceptional_mod)$r.squared,
                         com_lin_r2=summary(lin_common_mod)$r.squared,
                         com_quad_r2=summary(quad_common_mod)$r.squared,
                         delta_com_r2=summary(quad_common_mod)$r.squared-summary(lin_common_mod)$r.squared,
                         Set = paste0('decile ',this_decile)))

  
}
```


```{r, echo = FALSE, message=FALSE, warning = FALSE}
# pivot longer to create t-test df
t_data = delta_r2_df_dec %>% 
  select(Set, delta_com_r2,delta_ex_r2) %>% 
  pivot_longer(cols=c('delta_com_r2','delta_ex_r2'),
               names_to='problem', values_to = 'deltaR2') 

t_data_wide = delta_r2_df_dec %>% 
  select(Set, delta_com_r2,delta_ex_r2)

# compare change in R2 across linear and quadratic decile models
t.test(t_data_wide$delta_com_r2,
       t_data_wide$delta_ex_r2,
       alternative = 'two.sided',
       var.equal = T,paired=T)

# Calculate effect size for difference in R2 across decile models
cohens_d(t_data_wide$delta_com_r2,
       t_data_wide$delta_ex_r2,
       alternative = 'two.sided',
       var.equal = F,paired=T)

```
# Supplemental Figures

## Supplemental Figure 1

```{r, echo = FALSE, message=FALSE, warning = FALSE}


full_df %>% 
  mutate(cohort = as.factor(cohort),
         `Problem Type` = if_else(class=='Standard','Common','Exceptional'),
         firstButtonRT = firstButtonRT*1000) %>%
  group_by(pid,cohort,`Problem Type`) %>% 
  dplyr::summarise(count = n(), 
            first_rt = mean(firstButtonRT,na.rm=T),
            total_rt = mean(rt,na.rm=T)) %>% 
  pivot_longer(cols=contains('_rt'),names_to = 'rt_type',values_to='rt')%>% 
  ungroup() %>% 
  group_by(cohort,rt_type,`Problem Type`) %>% 
  summarise(cohort_mean_rt = mean(rt, na.rm=T),
            std_dev = sd(rt,na.rm=T),
            count = n())%>% 
  mutate(se = std_dev/sqrt(count),
         lower.ci = cohort_mean_rt - qt(1 - (0.05 / 2), count - 1) * se,
         upper.ci = cohort_mean_rt + qt(1 - (0.05 / 2), count - 1) * se)%>% 
  mutate(cohort = case_when(cohort == 0~"3",
                            cohort == 1~"5",
                            cohort == 2~"7"),
         rt_type = case_when(rt_type=='first_rt'~'First Button Press',
                             rt_type=='total_rt'~'Submit Press'))%>% 
  ggplot(aes(x = cohort, y = cohort_mean_rt,color = rt_type,group=rt_type))+
  geom_point()+
  geom_line()+
  geom_ribbon(aes(ymin = lower.ci, ymax= upper.ci, fill = rt_type), alpha = 0.2,colour = NA)+
  ylab("Time (ms)")+
  labs(color = "Reaction Time Type", fill = "Reaction Time Type")+
  xlab("Grade")+
  theme_classic()+
  facet_grid(.~`Problem Type`)

```


## Supplemental Figure 2

```{r s2, echo = FALSE, message=FALSE, warning = FALSE}


probsize_op_plot_rt = full_df_ses %>%
  # filter(timepoint==0) %>%
    filter(class=='Standard') %>%
    filter(!is.na(class)) %>% 
    filter(module=='MATH_FLUENCY') %>%
  # filter(operand_distance!=8) %>% 
  mutate(operation_type=if_else(operation_type == 'addition',
                                "Addition","Subtraction")) %>%
  group_by(pid, cohort, problem_size,low_income,operation_type) %>%
  dplyr::summarise(count = n(),
            sum_corr = sum(answered_correctly, na.rm = T),
            rt = mean(firstButtonRT, na.rm = T),
            rcs = (60*sum(answered_correctly, na.rm = T))/sum(firstButtonRT, na.rm = T)) %>%
  ungroup() %>%
  mutate(cohort=as.factor(cohort+3+cohort)) %>% 
  ggplot(aes(x = problem_size, y = rt,
             color=cohort, fill=cohort))+
  geom_smooth(method='lm')+
  ylab("Reaction Time (s)")+
  xlab("Set Size")+
  labs(color = "Grade", fill = 'Grade')+
  # ggtitle("RPM by Set Size")+
  theme_classic()+
  facet_wrap(.~operation_type, scale='free_x')

problem_type_plot_rt = full_df_ses %>%
  # rename(`Problem Type` = fluency_type) %>% 
  mutate(`Problem Type` = if_else(class=='Standard','Common','Exceptional')) %>%
    filter(!is.na(class)) %>% 
    filter(module=='MATH_FLUENCY') %>%
  # filter(operand_distance!=8) %>% 
  group_by(pid, grade,`Problem Type`) %>%
  dplyr::summarise(count = n(),
            sum_corr = sum(answered_correctly, na.rm = T),
            rt = mean(firstButtonRT, na.rm = T),
            rcs = (60*sum(answered_correctly, na.rm = T))/sum(firstButtonRT, na.rm = T)) %>%
  ungroup() %>%
  mutate(accuracy=sum_corr/count) %>%
  group_by(grade,`Problem Type`) %>%
  summarise(count=n(),
            mean_rt = mean(rt, na.rm=T),
            sd_rt = sd(rt, na.rm=T)) %>% 
      mutate(se = sd_rt/sqrt(count),
             lower.ci = mean_rt - qt(1 - (0.05 / 2), count - 1) * se,
             upper.ci = mean_rt + qt(1 - (0.05 / 2), count - 1) * se)%>%
  ggplot(aes(x = grade, y = mean_rt))+
  geom_point(aes(color = `Problem Type`))+
  geom_line(aes(color = `Problem Type`))+
  geom_ribbon(aes(ymin=lower.ci,ymax=upper.ci,fill=`Problem Type`),alpha=0.25)+
  ylab("Reaction Time (s)")+
  xlab("Grade")+
  labs(color = "Problem Set", fill = "Problem Set")+
  # ggtitle("RPM by Problem Type")+
  theme_classic()



((probsize_op_plot_rt|problem_type_plot_rt))& theme(legend.position = 'bottom')

```

### Supplemental Figure 3

```{r s3, echo = FALSE, message=FALSE, warning = FALSE}

operation_probsize_plot_acc = full_df_ses %>%
  # filter(timepoint==0) %>% 
  filter(!is.na(class)) %>% 
  # filter(class=='Standard') %>% 
  filter(module=='MATH_FLUENCY') %>% 
  mutate(operation_type=if_else(operation_type == 'addition',
                                "Addition","Subtraction"),
         cohort=as.factor(cohort+3+cohort)) %>%
  group_by(pid, cohort, problem_size,low_income,operation_type) %>%
  dplyr::summarise(count = n(),
            acc = sum(answered_correctly, na.rm = T)/n(),
            rt = mean(firstButtonRT, na.rm = T),
            rcs = (60*sum(answered_correctly, na.rm = T))/sum(firstButtonRT, na.rm = T)) %>%
  ungroup() %>% 
  ggplot(aes(x = problem_size, y = acc, group=cohort,
             color = cohort, fill= cohort))+
  geom_smooth(method='lm')+
  ylab("Accuracy")+
  xlab("Operation")+
  labs(color = "Grade", fill = "Grade")+
  # ggtitle("RPM by Operation Type")+
  theme_classic()+
  facet_grid(.~operation_type, scale='free_x')


problem_type_plot_acc = full_df %>%
  # rename(`Problem Type` = fluency_type) %>% 
  mutate(`Problem Type` = if_else(class=='Standard','Common','Exceptional')) %>%
    filter(!is.na(class)) %>% 
    filter(module=='MATH_FLUENCY') %>%
  # filter(operand_distance!=8) %>% 
  group_by(pid, grade,`Problem Type`) %>%
  dplyr::summarise(count = n(),
            sum_corr = sum(answered_correctly, na.rm = T),
            rt = mean(firstButtonRT, na.rm = T),
            rcs = (60*sum(answered_correctly, na.rm = T))/sum(firstButtonRT, na.rm = T)) %>%
  ungroup() %>%
  mutate(accuracy=sum_corr/count) %>%
  group_by(grade,`Problem Type`) %>%
  summarise(count=n(),
            mean_acc = mean(accuracy, na.rm=T),
            sd_acc = sd(accuracy, na.rm=T)) %>% 
      mutate(se = sd_acc/sqrt(count),
             lower.ci = mean_acc - qt(1 - (0.05 / 2), count - 1) * se,
             upper.ci = mean_acc + qt(1 - (0.05 / 2), count - 1) * se)%>%
  ggplot(aes(x = grade, y = mean_acc))+
  geom_point(aes(color = `Problem Type`))+
  geom_line(aes(color = `Problem Type`))+
  geom_ribbon(aes(ymin=lower.ci,ymax=upper.ci,fill=`Problem Type`),alpha=0.25)+
  ylab("Accuracy")+
  xlab("Grade")+
  labs(color = "Problem Set", fill = "Problem Set")+
  # ggtitle("RPM by Problem Type")+
  theme_classic()



((operation_probsize_plot_acc|problem_type_plot_acc))& theme(legend.position = 'bottom')


```


### Supplemental Figure 4
 
```{r, echo = FALSE, message=FALSE, warning = FALSE}



delta_r2_df %>% 
  # filter(str_contains(Set, 'decile')==F) %>% 
  select(Set, delta_com_r2,delta_ex_r2) %>% 
  pivot_longer(cols=c('delta_com_r2','delta_ex_r2'),
               names_to='Problem Set', values_to = 'Delta R2') %>% 
  mutate(`Problem Set` = case_when(`Problem Set` == 'delta_com_r2' ~ 'Common',
                                   `Problem Set` == 'delta_ex_r2' ~ 'Exceptional'),
         Set = case_when(Set == 'Quartile 1' ~ "SBAC Quartile: Lowest",
                         Set == 'Quartile 2' ~ "SBAC Quartile: Low",
                         Set == 'Quartile 3' ~ "SBAC Quartile: High",
                         Set == 'Quartile 4' ~ "SBAC Quartile: Highest",
                         T~Set)) %>% 
  mutate(Set = factor(Set, levels = c("Full", "Cohort 0", "Cohort 1", "Cohort 2", 
                                      "SBAC Quartile: Lowest","SBAC Quartile: Low",
                                      "SBAC Quartile: High", "SBAC Quartile: Highest"))) %>% drop_na() %>% 
  ggplot(aes(x=Set, y = `Delta R2`,fill=`Problem Set`))+
  geom_col(position='dodge')+
  theme_bw()+
   theme(axis.text.x = element_text(angle = 45, hjust=1))+
  ggtitle("Change in R2 with the Addition of Quadratic Term")


```

### Supplemental Table 1

# Examine Multiplication vs Addition/Subtraction Completion

```{r, echo = FALSE, message=FALSE, warning = FALSE}

# Load Raw Data
t0_sea <- read_csv("/Users/Ethan/Documents/Stanford/EdNeuro/UCSF/most updated data/t1_SEA_All_correctednamesdemo_rmdups_2020-02-28.csv",
                   col_types = cols(rt = col_character(),
                                    grade = col_character(),
                                    time = col_character(),
                                    trial_onset = col_character()))
# 977 completed math fluency
t0_sea %>% 
  filter(module=='MATH_FLUENCY') %>% 
  select(pid, question_text, operation_type, question_id) %>% 
  group_by(pid, operation_type) %>% 
  summarise(count=n()) %>% 
  ungroup() %>% 
  group_by(operation_type) %>% 
  mutate(number_ind_response = n()) %>% 
  ungroup() %>% 
  select(operation_type, number_ind_response) %>% 
  unique()%>% 
  mutate(perc = number_ind_response/977) 


mult_df = t0_sea %>% 
  filter(module=='MATH_FLUENCY') %>% 
  select(pid, operation_type) %>% 
  unique() %>% 
  group_by(pid) %>% 
  summarise(count=n()) %>% 
  mutate(did_mult = if_else(count==3, 1, 0)) %>% 
  select(pid, did_mult)
  

t0_sea %>% 
  filter(module=='MATH_FLUENCY') %>% 
  mutate(fluency_type = case_when(str_detect(question_text, "\\+ 10") ~ "Common",      
                                  str_detect(question_text, "\\- 1") ~ "Exceptional",          
                                  str_detect(question_text, "\\+ 1") ~ "Exceptional",
                                  str_detect(question_text, "\\- 0") ~ "Exceptional",
                                  str_detect(question_text, "\\+ 0") ~ "Exceptional",
                                  str_detect(question_text, "1 \\+") ~ "Exceptional",
                                  str_detect(question_text, "1 \\-") ~ "Exceptional",
                                  str_detect(question_text, "10 \\+") ~ "Common",
                                  str_detect(question_text, "0 \\+") ~ "Exceptional",
                                  str_detect(question_text, "1 \\+ 1") ~ "Exceptional",
                                  str_detect(question_text, "1 \\- 1") ~ "Exceptional",
                                  str_detect(question_text, "2 \\+ 2") ~ "Exceptional",
                                  str_detect(question_text, "2 \\- 2") ~ "Exceptional",
                                  str_detect(question_text, "3 \\+ 3") ~ "Exceptional",
                                  str_detect(question_text, "3 \\- 3") ~ "Exceptional",
                                  str_detect(question_text, "4 \\+ 4") ~ "Exceptional",
                                  str_detect(question_text, "4 \\- 4") ~ "Exceptional",
                                  str_detect(question_text, "5 \\+ 5") ~ "Exceptional",
                                  str_detect(question_text, "5 \\- 5") ~ "Exceptional",
                                  str_detect(question_text, "6 \\+ 6") ~ "Exceptional",
                                  str_detect(question_text, "6 \\- 6") ~ "Exceptional",
                                  str_detect(question_text, "7 \\+ 7") ~ "Exceptional",
                                  str_detect(question_text, "7 \\- 7") ~ "Exceptional",
                                  str_detect(question_text, "8 \\+ 8") ~ "Exceptional",
                                  str_detect(question_text, "8 \\- 8") ~ "Exceptional",
                                  str_detect(question_text, "9 \\+ 9") ~ "Exceptional",
                                  str_detect(question_text, "9 \\- 9") ~ "Exceptional",
                                  str_detect(question_text, "\\* 1") ~ "Filler",
                                  str_detect(question_text, "1 \\*") ~ "Filler",
                                  str_detect(question_text, "\\* 0") ~ "Filler",
                                  str_detect(question_text, "0 \\*") ~ "Filler",
                                  str_detect(question_text, "x") ~ "Filler",
                                  TRUE ~ "Common"))%>% 
  select(question_text, fluency_type)%>% 
  group_by(question_text,fluency_type) %>% 
  summarise(count=n())%>% 
  arrange(desc(fluency_type),question_text)


```

### Analysis for Supplemental Text

## Set Size*Operation -- Full Sample -- Supplement
```{r supplemental analyses, echo = FALSE, message=FALSE, warning = FALSE}

setsize_plot = full_df %>%
  group_by(pid, cohort, problem_size,operation_type) %>%
  dplyr::summarise(count = n(),
            sum_corr = sum(answered_correctly, na.rm = T),
            mean_rt = mean(firstButtonRT, na.rm = T),
            rcs = (60*sum(answered_correctly, na.rm = T))/sum(firstButtonRT, na.rm = T)) %>%
  ungroup() %>%
  mutate(cohort=as.factor(cohort+3+cohort)) %>% 
  ggplot(aes(x = problem_size, y = rcs,
             color=cohort,fill = cohort, group=cohort))+
  geom_smooth(method='lm')+
  ylab("Correct Responses per Minute")+
  xlab("Set Size")+
  labs(color = "Grade", fill = "Grade")+
  theme_classic()+
  facet_grid(.~operation_type, scales='free_x')+
  ylim(19,42)

setsize_mod_data = full_df %>%
    filter(module=='MATH_FLUENCY') %>% 
  group_by(pid, cohort, problem_size,operation_type) %>%
  dplyr::summarise(count = n(),
            sum_corr = sum(answered_correctly, na.rm = T),
            mean_rt = mean(firstButtonRT, na.rm = T),
            rcs = (60*sum(answered_correctly, na.rm = T))/sum(firstButtonRT, na.rm = T)) %>% 
  ungroup() %>% 
  unique()


setsize_mod_full = lmer(rcs ~ problem_size*operation_type*cohort+(1|pid), data = setsize_mod_data)

summary(setsize_mod_full)

effect_size_calc(setsize_mod_full) %>% 
  select(Effect, Rsq)

```
### Problem Type -- Full Sample  -- Supplement
```{r problem_type_supp, echo = FALSE, message=FALSE, warning = FALSE}
problem_type_plot = full_df %>%
  mutate(`Problem Type` = if_else(class=='Standard','Common','Exceptional')) %>%
    filter(!is.na(class)) %>% 
    filter(module=='MATH_FLUENCY') %>%
  group_by(pid, grade,`Problem Type`) %>%
  dplyr::summarise(count = n(),
            sum_corr = sum(answered_correctly, na.rm = T),
            mean_rt = mean(firstButtonRT, na.rm = T),
            rcs = (60*sum(answered_correctly, na.rm = T))/sum(firstButtonRT, na.rm = T)) %>%
  ungroup() %>%
  mutate(accuracy=sum_corr/count) %>%
  group_by(grade,`Problem Type`) %>%
  summarise(count=n(),
            mean_rpm = mean(rcs, na.rm=T),
            sd_rpm = sd(rcs, na.rm=T)) %>% 
      mutate(se = sd_rpm/sqrt(count),
             lower.ci = mean_rpm - qt(1 - (0.05 / 2), count - 1) * se,
             upper.ci = mean_rpm + qt(1 - (0.05 / 2), count - 1) * se)%>%
  ggplot(aes(x = grade, y = mean_rpm))+
  geom_point(aes(color = `Problem Type`))+
  geom_line(aes(color = `Problem Type`))+
  geom_ribbon(aes(ymin=lower.ci,ymax=upper.ci,fill=`Problem Type`),alpha=0.25)+
  ylab("Correct Responses per Minute")+
  xlab("Grade")+
  labs(color = "Problem Set", fill = "Problem Set")+
  # ggtitle("RPM by Problem Type")+
  theme_classic()+
  ylim(19,42)

mod_data = full_df %>%
    mutate(ProblemType = if_else(class=='Standard','Standard','Non-Standard')) %>%
  # filter(timepoint==0) %>%
    # filter(class=='Standard') %>% 
    filter(!is.na(class)) %>% 
    filter(module=='MATH_FLUENCY') %>%
  # filter(operand_distance!=8) %>% 
  mutate(operand_distance=as.factor(operand_distance)) %>% 
  group_by(pid, cohort, ProblemType) %>%
  dplyr::summarise(count = n(),
            sum_corr = sum(answered_correctly, na.rm = T),
            mean_rt = mean(firstButtonRT, na.rm = T),
            rcs = (60*sum(answered_correctly, na.rm = T))/sum(firstButtonRT, na.rm = T)) %>% 
  ungroup() 

probTypeMod_full = lmer(rcs ~ ProblemType*cohort + (1|pid), data = mod_data)

summary(probTypeMod_full)

effect_size_calc(probTypeMod_full) %>% 
  select(Effect,Rsq)

```

## Corrected p-values

```{r, echo = FALSE, message=FALSE, warning = FALSE}

## Get FDR Corrected p-values for effects

probSum = as.data.frame(summary(probTypeMod_full)$coefficients) %>% 
  mutate(model="Problem Type")

setSizeSum = as.data.frame(summary(setsize_mod_full)$coefficients) %>% 
  mutate(model="SetSize")

# Print adjusted p-values
probSum %>% 
  rbind(setSizeSum) %>% 
  mutate(p.adj_fdr = p.adjust(`Pr(>|t|)`,"fdr"),
         p.adj_bonf = p.adjust(`Pr(>|t|)`,"bonf")) %>% 
  select(model, p.adj_fdr,p.adj_bonf)


```
